Epistemic Semantics in Guarded String Models

### 2020-09-22 tex/sdraft.tex
figure hacking

\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
\input{syntax.tex}
\usepackage{wasysym,graphicx}
\usepackage{tcolorbox}

@@

\begin{figure}
  \[\begin{array}{lcl}
  \multicolumn 3 l {\textrm{state formulas}\hfill (a \in \mathsf{B}) \hfill \mbox{}}\\
  \rho,\sigma,\varphi &::=& a \mid0 \mid 1 \mid\rho+\sigma \mid \rho\,\sigma \mid \bar{\rho} \\ \\
  \multicolumn {3} l {\textrm{effect formulas}}\\
  \zeta,\eta & ::= & \rho \mathbin{:} \sigma \mid \zeta+\eta
    \mid \zeta \mathbin{\&} \eta \mid \bar{\zeta}
  \end{array}\]

\[
\begin{array}{lll}
  \sem{\rho\mathbin{:}\sigma}^\varphi & = &
  {\mathcal A}^{\rho\varphi}_{\mathsf{B}}\times{\mathcal A}^{\sigma\varphi}_{\mathsf{B}}\\[.1em]
\sem{\zeta +\eta}^\varphi & = & \sem{\zeta}^\varphi\cup\sem{\eta}^\varphi\\
\sem{\zeta \mathbin{\&}  \eta}^\varphi & = & \sem{\zeta}^\varphi\cap\sem{\eta}^\varphi\\
\sem{\bar{\zeta}}^\varphi & = & {\mathcal A}^{\varphi}_{\mathsf{B}} \times {\mathcal A}^{\varphi}_{\mathsf{B}} \setminus \sem{\zeta}^\varphi\\
\end{array}  
\]

\vspace{-2mm}

\caption{Syntax of state formulas and syntax and semantics of effect formulas. Effect formulas
  denote relations between atoms. In a state formula, juxtaposition $\rho\,\sigma$ is conjunction.}
\label{effectfigure}
\end{figure}

@@

Epik is a specification language for possible worlds models that
includes declarations of events and states, state formulas, effect
formulas, and additional information.  Figure \ref{epikfigure} shows
an Epik program that describes a possible worlds model for two agents
with information about one coin, and events of the agents
semi-privately looking at the coin. The line beginning with {\tt
  state} enumerates ${\sf{B}}$. The line beginning with {\tt
  constraint} gives the state formula.  The lines beginning with {\tt
  event} declare events and their effect formulas.  Finally the lines
beginning with {\tt agent} define {\em event alternative} relations
for agents.  Each clause with an arrow has a single event symbol on
the left, and a disjunction of alternative events on the right of the
arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks
at heads), is that when $b_1$ happens, for Amy either $b_1$ or $b_0$
(Bill peeks at tails) could be happening.

\begin{figure}
  \footnotesize
\begin{center}
{\tt
\begin{tabular}{l}
state h t \\
constraint h!t + t!h \\
event a1 h:h \\
event a0 t:t \\
event b1 h:h \\
event b0 t:t \\
agent aly \\
\hspace{3mm} a1 -> a1 \\
\hspace{3mm} a0 -> a0 \\
\hspace{3mm} b1 -> b1 + b0\\
\hspace{3mm} b0 -> b1 + b0\\
agent bob \\
\hspace{3mm} b1 -> b1 \\
\hspace{3mm} b0 -> b0 \\
\hspace{3mm} a1 -> a1 + a0\\
\hspace{3mm} a0 -> a1 + a0\\
\end{tabular} }
\end{center}

@@

\begin{figure}
\begin{tabular}{ll}
events & $e \in \mathsf{P}$ \\
states & $\sigma$ as in Figure 1 \\    
$p,q$ & $::= e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p$\\
\end{tabular}

\vspace{1mm}
\begin{tabular}{cl}
$\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
\(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
$\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
\end{tabular}

@@

\begin{figure}
  \centering
  \(\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ b }^{\epikspec} && \mathcal A^{b\varphi}_{\mathsf B}\\
  \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
  \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
  \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
  \sem{p^*}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
  \sem{ \neg p }^{\epikspec} && \sem{\bullet^*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
  \end{array}
  \)

\hspace{2mm}\(\sem{ \Diamond_a\,p } \hspace{4mm}{\triangleq} \hspace{4mm}\{x \mid \exists y . x{\hat{R}_a}y \wedge y \in \sem{ p }^{\epikspec}\} \)
  \caption{Interpretation of Epik terms as sets of guarded strings}
  \label{fig:interp2}
\end{figure}

@@

\begin{figure}
\footnotesize
\begin{hangingpar}
{\tt St} \hspace{2mm} Tests such as 0 1 1 0. The length is the number of generators.
\end{hangingpar}

\begin{hangingpar}
{\tt UnequalStPair} \hspace{2mm} Sequence of two unequal tests such as 0 1 1 0 0 1 1 1,
    differing in one or more positions.
\end{hangingpar}

@@

\begin{hangingpar}
\begin{verbatim}
define Kst(X) St |  Kpl(X);
\end{verbatim}
KAT Kleene plus and Kleene star. The Fst operation {\tt |} is union.
\end{hangingpar}

\caption{\raggedright Definition in Fst of KAT product and KAT Kleene star.
  Where {\tt X} and {\tt Y} are regular sets and
  {\tt R} and {\tt S} are regular relatiions,
  {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
  {\tt X|Y} is the union of {\tt X} and {\tt Y},
  {\tt \~{}X} is the complement of {\tt X},
  {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
  {\tt R{.}l} is the co-domain of {\tt R}, and
  {\tt {\$}X} is the set of strings that have a substring in $X$.   
}
\label{fstfigure1}
\end{figure}

@@

\begin{figure} \label{relclosure}
  \footnotesize
{\raggedright
  \begin{hangingpar}
{\tt{define RelKpl(R)}}\\
$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
$\underbrace{\mbox{\tt{[R+]}}}_a$
$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
\end{hangingpar}}

@@

\begin{hangingpar}
{\tt{define Kst(R)
    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}\\ The Fst
operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
{\tt{R}}.
\end{hangingpar}

\caption{Definition in Fst of the Kleene concatenation closure of
  a relation between guarded strings.}
\label{fstfigure2}
\end{figure}

@@

\begin{figure}
  \centering
  \(\begin{array}{r >{\triangleq}c l}
  \semlst p 0 && [] \\
  \semlst 0 n && [] \\
  \semlst 1 n && \setToList \AphiB \\
  \semlst e n && [ \alpha e \beta \mid \alpha \mathrel{\zeta_e} \beta ] \\
  \semlst b n && \setToList{ \mathcal A_{\mathsf B}^{b\psi}} \\
  \semlst{p + q} n && \semlst p n \mathbin{\texttt{+\!+}} \semlst q n \\
  \semlst{p;q} n &&
  \left(\semlst p n \diamond \semlst q n \right)|_n\\
  \semlst{p^*} n && [] + \left(\semlst p n \diamond \semlst{p^*} {n-i}\right) |_n \\
  \multicolumn 3 r {\text{where}~
    \begin{array}[t]{r >{\hspace{-.7em}}l}
      i &= \max\{1,\min\{|g|\mid g \in \semlst p n \}\} \\
    \end{array}}\\
  \semlst{\neg p} n && \semlst {\bullet^*} n \setminus \semlst p n \\
  \semlst{\Diamond_a~p } n &&
         [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\semlst p n]
  \end{array}
  \)
  \caption{Bounded interpretation using lazy lists}
  \label{fig:bound-interp}
\end{figure}

@@

\begin{figure}
  \footnotesize
\begin{tabular}[t]{l|c|l}
Amy & $e$ & ${R}_a$\\
Bob & $e$ & ${R}_b$\\
it & $d$ & $d$ \\
heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
& $\lambda p~m~R.{\neg}(m\;({\neg}p)\;R)$ \\
whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
& $\lambda p~m~R.{\neg}(m\;({\neg}p)\;R)$ \\
&& \multicolumn{1}{r}{$+ {\neg}(m\;p\;R)$} \\
\end{tabular}
\caption{Partial categorial grammar lexicon. The first column has a word form.
  the second column a categorial type, and third column a semantic translation
  in a logical language that extends the Epik term language with lambda. }
\label{lexiconfigure}
\end{figure}

@@

\bibliography{anthology,acl2020}
\bibliographystyle{acl_natbib}

>>>>

\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
\input{syntax.tex}
\usepackage{wasysym,graphicx}
\usepackage{tcolorbox}
\usepackage{booktabs}

@@

\begin{figure}
  \fbox{
    \begin{minipage}{\columnwidth}
  \(\begin{array}{lcl}
  \multicolumn 3 l {\textrm{state formulas}\hfill (a \in \mathsf{B}) \hfill \mbox{}}\\
  \rho,\sigma,\varphi &::=& a \mid0 \mid 1 \mid\rho+\sigma \mid \rho\,\sigma \mid \bar{\rho} \\
  \multicolumn {3} l {\textrm{effect formulas}}\\
  \zeta,\eta & ::= & \rho \mathbin{:} \sigma \mid \zeta+\eta
    \mid \zeta \mathbin{\&} \eta \mid \bar{\zeta}
  \end{array}\)

  \(
  \begin{array}{lll}
    \sem{\rho\mathbin{:}\sigma}^\varphi & = &
        {\mathcal A}^{\rho\varphi}_{\mathsf{B}}\times{\mathcal A}^{\sigma\varphi}_{\mathsf{B}}\\[.1em]
        \sem{\zeta +\eta}^\varphi & = & \sem{\zeta}^\varphi\cup\sem{\eta}^\varphi\\
        \sem{\zeta \mathbin{\&}  \eta}^\varphi & = & \sem{\zeta}^\varphi\cap\sem{\eta}^\varphi\\
        \sem{\bar{\zeta}}^\varphi & = & {\mathcal A}^{\varphi}_{\mathsf{B}} \times {\mathcal A}^{\varphi}_{\mathsf{B}} \setminus \sem{\zeta}^\varphi\\
  \end{array}  
  \)
  \end{minipage}}
\caption{Syntax of state formulas and syntax and semantics of effect formulas. Effect formulas
  denote relations between atoms. In a state formula, juxtaposition $\rho\,\sigma$ is conjunction.}
\label{effectfigure}
\end{figure}

@@

Epik is a specification language for possible worlds models that
includes declarations of events and states, state formulas, effect
formulas, and additional information.  Figure \ref{epikfigure} shows
an Epik program that describes a possible worlds model for two agents
with information about one coin, and events of the agents
semi-privately looking at the coin. The line beginning with {\tt
  state} enumerates ${\sf{B}}$. The line beginning with {\tt
  restrict} gives the state formula.  The lines beginning with {\tt
  event} declare events and their effect formulas.  Finally the lines
beginning with {\tt agent} define {\em event alternative} relations
for agents.  Each clause with an arrow has a single event symbol on
the left, and a disjunction of alternative events on the right of the
arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks
at heads), is that when $b_1$ happens, for Amy either $b_1$ or $b_0$
(Bill peeks at tails) could be happening.

\begin{figure}
  \begin{tcolorbox}
    \footnotesize
    \begin{minipage}{0.45\textwidth}
      \begin{center}    
        {\tt
          \begin{tabular}{l}
            state h t \\
            restrict h!t + t!h \\
            event a1 h:h \\
            event a0 t:t \\
            event b1 h:h \\
            event b0 t:t \\
      \end{tabular}}
      \end{center}
    \end{minipage} \hfill 
    \begin{minipage}{0.45\textwidth}
      \begin{center}
        {\tt
          \begin{tabular}{l}
            agent aly \\
            \hspace{3mm} a1 -> a1 \\
            \hspace{3mm} a0 -> a0 \\
            \hspace{3mm} b1 -> b1 + b0\\
            \hspace{3mm} b0 -> b1 + b0\\
            agent bob \\
            \hspace{3mm} b1 -> b1 \\
            \hspace{3mm} b0 -> b0 \\
            \hspace{3mm} a1 -> a1 + a0\\
            \hspace{3mm} a0 -> a1 + a0\\
        \end{tabular} }
      \end{center}
    \end{minipage}
  \end{tcolorbox}

@@

\begin{figure}
  \fbox{
    \begin{minipage}{\columnwidth}
      \(\begin{array}{ll}
      \text{events} & e \in \mathsf{P} \\
      \text{states} & \sigma~\text{as in Figure 1} \\    
      p,q & ::= e\,|\,\sigma\,|\,p+q\,|\,pq\,|\,p^*\,|\,\neg p\,|\, \Diamond_ap \\
      \end{array}\)

      \vspace{1mm}
      \begin{tabular}{cl}
        $\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
        \(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
        $\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
      \end{tabular}
  \end{minipage}}

@@

\begin{figure}
  \fbox{\begin{minipage}{\columnwidth}
  \centering
  \(\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ b }^{\epikspec} && \mathcal A^{b\varphi}_{\mathsf B}\\
  \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
  \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
  \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
  \sem{p^*}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
  \sem{ \neg p }^{\epikspec} && \sem{\bullet^*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
  \end{array}
  \)

  \hspace{2mm}\(\sem{ \Diamond_a\,p } \hspace{4mm}{\triangleq} \hspace{4mm}\{x \mid \exists y . x{\hat{R}_a}y \wedge y \in \sem{ p }^{\epikspec}\} \)
  \end{minipage}}
  \caption{Interpretation of Epik terms as sets of guarded strings}
  \label{fig:interp2}
\end{figure}

@@

\begin{figure}
  \begin{tcolorbox}[size=fbox]
\footnotesize
\begin{hangingpar}
  {\tt St} \hspace{2mm} \\
  Atomic Tests such as $0110$.% The length is the number of generators.
\end{hangingpar}

\begin{hangingpar}
  {\tt UnequalStPair} \hspace{2mm} \\
  Sequence of two unequal tests such as $0110~0111$.
\end{hangingpar}

@@

\begin{hangingpar}
\begin{verbatim}
define Kst(X) St | Kpl(X);
\end{verbatim}
KAT Kleene plus and Kleene star. The Fst operation {\tt |} is union.
\end{hangingpar}
\end{tcolorbox}
\caption{\raggedright Definition in Fst of KAT product and KAT Kleene star.
  Where {\tt X} and {\tt Y} are regular sets and
  {\tt R} and {\tt S} are regular relatiions,
  {\tt X{\&}Y} is the intersection of {\tt X} and {\tt Y},
  {\tt X|Y} is the union of {\tt X} and {\tt Y},
  {\tt \~{}X} is the complement of {\tt X},
  {\tt R{.o.}S} is the composition of {\tt R} and {\tt S},
  {\tt R{.}l} is the co-domain of {\tt R}, and
  {\tt {\$}X} is the set of strings that have a substring in $X$.   
}
\label{fstfigure1}
\end{figure}

@@

\begin{figure} \label{relclosure}
  \begin{tcolorbox}[size=fbox]
  \footnotesize  
{\raggedright
  \begin{hangingpar}
%% {\tt{define RelKpl(R)}}\\
%% $\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
%% $\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
%% $\underbrace{\mbox{\tt{[R+]}}}_a$
%% $\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
%% $\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
{\tt{define RelKpl(R)}}\\
$\!\underbrace{\texttt{Squash.i}\,\texttt{.o.}}_c
\underbrace{\mbox{\tt Wf0\,.o.}}_b
\underbrace{\mbox{\tt{[R+]}}}_a
\underbrace{\mbox{\tt .o.\,Wf0}}_b
\underbrace{\mbox{\tt{.o.\,Squash}}}_c$

\end{hangingpar}}

@@

\begin{hangingpar}
{\tt{define Kst(R)
    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}\\ The Fst
operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
{\tt{R}}.
\end{hangingpar}
\end{tcolorbox}
\caption{Definition in Fst of the Kleene concatenation closure of
  a relation between guarded strings.}
\label{fstfigure2}
\end{figure}

@@

\begin{figure}[!t]
  \fbox{
    \begin{minipage}{\columnwidth}
      \hspace{-0.9em}\(\begin{array}{r >{\triangleq}c l}
      \semlst p 0 && [] \\
      \semlst 0 n && [] \\
      \semlst 1 n && \setToList \AphiB \\
      \semlst e n && [ \alpha e \beta \mid \alpha \mathrel{\zeta_e} \beta ] \\
      \semlst b n && \setToList{ \mathcal A_{\mathsf B}^{b\psi}} \\
      \semlst{p\! + \!q} n && \semlst p n \mathbin{\texttt{+\!+}} \semlst q n \\
      \semlst{p;q} n &&
      (\semlst p n \diamond \semlst q n)|_n\\
      \semlst{p^*} n && [] + (\semlst p n \diamond \semlst{p^*} {n-i}) |_n \\
      \multicolumn 3 r {\footnotesize \text{where}~
        \begin{array}[t]{r >{\hspace{-.7em}}l}
          i &= \max\{1,\min\{|g|\mid g \in \semlst p n \}\} \\
      \end{array}}\\
      \semlst{\neg p} n && \semlst {\bullet^*} n \setminus \semlst p n \\
      \semlst{\Diamond_ap } n &&
             [g' \mid g'\,\hat{R_a}\,g,\texttt{\footnotesize for}\,g\,\texttt{\footnotesize in}\,\semlst p n]
      \end{array}
      \)
    \end{minipage}}
  \caption{Bounded interpretation using lazy lists}
  \label{fig:bound-interp}
\end{figure}

@@

\begin{figure}[!t]
  \vspace{1em}
  \centering
  \footnotesize
  \begin{tabular}[t]{lcc}
    \toprule
        \textsc{Item} & \textsc{Type} & \textsc{Semantics}\\ \midrule  
        Amy & $e$ & ${R}_a$\\
        Bob & $e$ & ${R}_b$\\
        it & $d$ & $d$ \\
        heads & $d \backslash_D t$ & $\lambda x.\bullet^*h$ \\
        tails & $d \backslash_D t$ & $\lambda x.\bullet^*t$ \\
        is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P~x.~P~x$ \\
        knows & $(e \backslash t)\slash_M t$ &  $\lambda p~R.~{\Diamond}_Rp$\\
        that  & $(((e \backslash t){\slash_M}t) \backslash (e \backslash t)) \slash t$
        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
        whether  & $(((e \backslash t){\slash_M} t) \backslash (e \backslash t)) \slash t$ 
        & $\lambda p\,m\,R.{\neg}(m\,({\neg}p)\,R)$ \\
        && \multicolumn{1}{r}{$+ {\neg}(m\,p\,R)$} \\ \bottomrule
      \end{tabular}
  \caption{Partial categorial grammar lexicon. The first column has a word form.
    the second column a categorial type, and third column a semantic translation
    in a logical language that extends the Epik term language with lambda. }
  \label{lexiconfigure}
\end{figure}

@@

\bibliography{anthology,acl2020}
\bibliographystyle{chicago}%{acl_natbib}
